In-silico molecular enrichment and clearance of the human intracranial space

### 2025-01-23 paper/supplementary_main.tex
Updates in said caption. Spell-check on supplementary.

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2024modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivacular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

@@

Finally, we impose a homogogenous Neumann condition at all network end
nodes to augment~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
    \label{eq:1d_bcs}
\end{equation}
This no-flux condition states that the network end nodes represent
barriers for molecular efflux into or out of the perivascular network,
and thus that all solute exchange between the PVSs and their
surroundings takes place via the lateral outer PVS surface and is
regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
and~\eqref{eq:1d_pde}. Larger particles have been observed to
accumulate within the PVS as the surface arteries penetrate into the
brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
and~\eqref{eq:1d_bcs} is appropriate to represent such
behavior. However, to represent a continuously extending PVSs also
along penetrating vessels, this condition should be revisited.

@@

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diamter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

 \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} comlemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:

@@

This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:results1}A, C,
\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
associated with the nearest supply node, and edges between nodes are
preserved. Next, we compute a minimal, bifurcating and directed tree
representation of each subnetwork: $\mathcal{T}^1, \mathcal{T}^2,
\mathcal{T}^3$. Each tree $\mathcal{T}^j = \cup E_n^j$ consists of the
subset of the nodes from $\Lambda^j$ that have degree 1 (are leaf or
root nodes) or degree $3$ (are true bifurcation points), and each path
between nodes with degree $\not = 2$ in $\Lambda^j$ is represented in
$\mathcal{T}^j$ by an edge $E^j$ with edge length $L$ corresponding to
the total length of the original path and edge radius $R^j$ as the
average of the path radii.

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional}. We next uniquely assign the flow rate
$\langle Q'_n \rangle$ to each of the branches $\Lambda_i$ that form
the path $E_n^j$, thus yielding $\langle Q'_i \rangle$ for each
perivascular segment $\Lambda_i$ (except those ignored above, for
which we set a flow rate of zero). Finally, we define the mean
longitudinal perivascular velocity induced by the peristaltic wave
$\langle \hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i
\rangle$ by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2)
= \pi (\beta_i^2 - 1) {R_1^i}^2$.

@@

We assess the accuracy and numerical convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions. Specifically, we generate a sequence of three meshes, see \Cref{fig:mesh_refinement}, with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest. The mesh refinement employed here is adapative in the sense that the coarser mesh is refined near expected sharp concentration gradients. This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.  
Similarly, we investigate the effect of the time step size by solving the intracranial transport model with varying time steps $dt \in \{60, 120, 240 \}$\,s.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/csf_minmax.png}
        \caption{Minimum and maximum tracer concentration in  the CSF under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1 \linewidth]{figures/par_minmax.png}
        \caption{Minimum and maximum tracer concentration in the parenchyma under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/art_minmax.png}
        \caption{Minimum and maximum tracer concentration in PVSs around arteries under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1 \linewidth]{figures/ven_minmax.png}
         \caption{Minimum and maximum tracer concentration in PVSs around veins under mesh refinement}
    \end{subfigure}
    \caption{Minumum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
    \label{fig:mesh_convergence_concentrations}
\end{figure}

>>>>

In this section, we describe the mathematical
model~\eqref{eq:multi_transport}, the associated definitions,
interface and boundary conditions in further detail. We refer to the
main text for explicit parameter values while giving the general,
abstract form here, and to the reference~\cite{masri2024modelling} for
the derivation and analysis of this 3D-1D model. Recall that we model
a concentration field $c = c(x, t)$ for $x \in \Omega$ and $t > 0$
defined in $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$ separately. In
each of these domains, $c$ satisfies
%\mer{MER: Rami asks if it should be a $\phi$ in the advective term here. We can include it, or drop it, but in the latter case, perhaps we should clarify that u is an average/superficial/Darcy velocity.}
\begin{equation}
  \partial_t (\phi c) - \nabla \cdot (D \nabla (\phi c) ) + \nabla \cdot (\bm u c ) + \xi (\overline{c} - \hat c ) \delta_\Gamma
  %= f
   = 0
  \quad \quad \mathrm{in} \quad \Omega_{\rm CSF}, \Omega_{\rm{PAR}}.
  \label{eq:3d_pde}
\end{equation}
In~\eqref{eq:3d_pde}, $\phi$ is the fluid volume fraction (also known as the porosity) defined in the parenchyma ($\phi \ll 1$ in $\Omega_{\rm PAR}$) and in the CSF spaces ($\phi = 1$ in $\Omega_{\rm CSF}$). In the parenchyma, $\phi$ represents the extracellular space volume fraction, and thus $c$ here generally represents the intrinsic (in contrast to the superficial) concentration~\cite{sykova2008diffusion}. Moreover, $D$ is the effective diffusion coefficient of the relevant solute in the respective media which takes different values over the CSF spaces and the parenchyma, depending on tortuosity~\cite{sykova2008diffusion} and dispersive effects; and $\bm u$ is a convective velocity field representing the flow of CSF in $\Omega_{\rm CSF}$ and the flow of ISF in $\Omega_{\rm PAR}$. $\xi$ models a transfer or exchange parameter between the 3D domain ($\Omega_{\rm PAR}$ or $\Omega_{\rm CSF}$) and the perivascular networks $\Lambda_a, \Lambda_v$. To summarize,
\begin{equation*}
\phi =  \begin{cases}
  1  & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  \phi_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
  \end{cases}, \; 
D = \begin{cases}
  D_{\rm{CSF}} & \mathrm{in} \;  \Omega_{\rm{CSF}} \\ 
  D_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} \end{cases}, \; 
  \quad \bm u  = \begin{cases}
  \bm{u}_{\rm{CSF}} & \mathrm{in} \; \Omega_{\rm{CSF}} \\ 
  \bm{u}_{\rm{PAR}} & \mathrm{in} \; \Omega_{\rm{PAR}} 
\end{cases}, \; 
\xi = \begin{cases}
  \xi_{\rm{CSF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{CSF}}| \neq 0 \\
  \xi_{\rm{EF}} & \mathrm{if} |\Omega^i \cap \Omega_{\rm{PAR}}| \neq 0  
\end{cases} . 
\end{equation*}
In the above, $\xi$ is defined segment-wise: for each centerline $\Lambda^i$ with surrounding PVS $\Omega^i$, $|\Omega^i \cap \Omega_{\rm{PAR}}|$ (resp. $|\Omega^i \cap \Omega_{\rm{CSF}}|$) is nonzero whenever $\Omega^i$ intersects $\Omega_{\rm{PAR}}$ (resp. $\Omega_{\rm{CSF}}$). If the surrounding PVS intersects both, then we set $\xi = \xi_{\rm EF}$ if $\Omega^i$ mainly ($80$ percent) intersects $\Omega_{\rm{PAR}}$, in which case the interaction with $\Omega_{\rm{CSF}}$ is ignored; otherwise, we set $\xi = \xi_{\rm CSF}$.

@@

Finally, we impose a homogeneous Neumann condition at all network end
nodes to augment~\eqref{eq:1d_pde}:
\begin{equation}
    \hat D A \partial_s (\phi \hat c) - A \hat u \hat c  = 0 \quad \mathrm{on} \quad \mathcal{N}.
    \label{eq:1d_bcs}
\end{equation}
This no-flux condition states that the network end nodes represent
barriers for molecular efflux into or out of the perivascular network,
and thus that all solute exchange between the PVSs and their
surroundings takes place via the lateral outer PVS surface and is
regulated by the exchange parameter $\xi$ cf.~\eqref{eq:3d_pde}
and~\eqref{eq:1d_pde}. Larger particles have been observed to
accumulate within the PVS as the surface arteries penetrate into the
brain parenchyma~\cite{mestre2018flow, bedussi2018paravascular},
and~\eqref{eq:1d_bcs} is appropriate to represent such
behavior. However, to represent a continuously extending PVSs also
along penetrating vessels, this condition should be revisited.

@@

We first consider the discretization of~\eqref{eq:3d_pde} and
introduce the discrete space
\begin{equation}
  %V_h^k = \{c_h \in L^2(\Omega): \;  v_h \vert_E \in \mathbb{P}^k(E), \; E \in \mathcal{T}_h\}.
  V_h = \{c \in L^2(\Omega): \;  c \vert_E \in \mathrm{P}^1(E), \; E \in \mathcal{T}\} ,
\end{equation}
where $L^2(\Omega)$ is the space of square-integrable functions
defined over $\Omega$ and $\mathrm{P}^1(E)$ denotes the space of
polynomials of total degree $\leqslant 1$ defined over the tetrahedra
$E$. To discretize the diffusion term in~\eqref{eq:3d_pde}, we use a
symmetric weighted interior penalty DG formulation, referring
to~\cite{ern2009discontinuous} and \cite[Section
  4.5.2.3]{di2011mathematical} for details on this
method. Recall that $\phi$ is constant in each domain
  $\Omega_{\rm PAR}$ and $\Omega_{\rm CSF}$, and thus in particular
  that $\nabla (\phi c) = \phi \nabla c$ on each $E \in \mathcal{T}$.
Define for $c, v \in V_h^k$:
\begin{equation}
  a_h (c, v)
  =
  \sum_{E \in \mathcal{T}} \int_{E} D\phi \nabla c \cdot \nabla v
  + \sum_{F \in \mathcal{F}_i} \eta \frac{\gamma_{D,F}}{h_F} \int_{F} [c][v] 
  - \sum_{F \in \mathcal{F}_i} \int_{F} \left( \{D \phi \nabla c \}_w \cdot \bm n_F [v]
  + \{D \phi \nabla v \}_w \cdot \bm n_F [c] \right).
  \label{eq:dg_sip}
\end{equation}
In~\eqref{eq:dg_sip}, for each facet $F \in \mathcal{F}^i$ shared
between cells $E^1$ and $E^2$, we associate a  facet normal vector $\bm n_F$
pointing from $E^1$ to $E^2$. The facet diameter is denoted by $h_F$,
the jump $[\cdot]$ is given by $[v] = v|_{E^1} - v|_{E^2}$, and the
unweighted average $\{ \cdot \}$ and weighed average $\{ \cdot \}_w$
are defined as:
\begin{equation}
  \{v\} = \frac12 v |_{E^1} + \frac12 v |_{E^2}, \qquad
  \{v\}_{w} = \frac{\kappa_2}{\kappa_1 + \kappa_2} v |_{E^1} + \frac{\kappa_1}{\kappa_1 + \kappa_2} v |_{E^2}
  \quad \text{where} \quad \kappa_i = (D \phi) |_{E^i}.  
\end{equation}
The parameter $\gamma_{D,F}$ is the harmonic mean of the
porosity-weighted diffusion coefficient given by
\begin{equation}
  \gamma_{D,F} = \frac{\kappa_1\kappa_2}{\kappa_1 + \kappa_2},
\end{equation}
and $\eta$ is a user-specified penalty parameter. To discretize the convection term in~\eqref{eq:3d_pde}, we use
upwinding, see \cite[Section 2.3.1]{di2011mathematical} and the
references therein. For $c, v \in V_h^k$, define
\begin{equation}
  a_h^{\mathrm{up}}(c, v)
  = - \sum_{E \in \mathcal{T}} \int_E \bm{u} c \nabla v
  + \sum_{F \in \mathcal{F}_i } \int_F \{ \bm u  c\} \cdot \bm{n}_F [v] +
  \sum_{F \in \mathcal{F}_i} \int_F \frac{|\bm u \cdot \bm n_F|}{2} [c] [v] .
\end{equation}
Our discrete formulation for \eqref{eq:3d_pde} with the interface conditions~\eqref{eq:3d_interface} and boundary conditions~\eqref{eq:3d_bcs} and given initial conditions $c_h^0$ then reads: for $n = 1, 2, \dots$, with $t^n - t^{n-1} = \tau$, find $c_h^n \in V_h$ such that for all $v \in V_h$:
\begin{equation} \label{eq:3d_dg_disc}
  \int_{\Omega} \frac{1}{\tau} (\phi c_h^n - \phi c_h^{n-1}) v
  + a_h(c_h^n, v) + a_h^{\mathrm{up}}(c_h^n,v)
  + \int_{\Gamma_{\rm pia} \cup \Gamma_{\rm LV}} \beta_{\rm pia}[c_h^n][v] 
  + \int_{\Gamma_{\rm AM-U}} \beta_{\rm exit} c_h^n v 
  + \int_{\Lambda} \xi  P \left (\overline{c_h^n} - \hat c_h^n \right) \overline{v} 
  = 
  %\int_{\Omega} f^n v +
  \int_{\Gamma_{\rm SSAS}} g_{\rm influx} v
  %, \quad \forall v \in V_h^k
  .
\end{equation}

@@

 \textit{Summary.} The discretization for \eqref{eq:3d_pde} and \eqref{eq:1d_pde} complemented with boundary conditions (\eqref{eq:3d_bcs} and \eqref{eq:1d_bcs}), interface and bifurcation  conditions (\eqref{eq:3d_interface}, \eqref{eq:1d_flow:bifurcations}, and \eqref{eq:conc_continuity}), and intial conditions $c_h^0 =0 $ and $\hat c_h^0 = 0 $ is the following:

@@

This theoretical formalism~\cite{gjerde2023directional} is defined
relative to a network in the form of a directed, bifurcating tree with
a single supply node/root $i_0$. To extend to a network of cerebral
arteries with multiple supply nodes (such as the basilar and two
internal carotid arteries in the current data
set~\cite{hodneland2019new}), we separate the network $\Lambda$ into
edge-disjoint subnetworks $\Lambda^1, \Lambda^2, \Lambda^3$, one for
each of the supply nodes (\Cref{fig:results1}A, C,
\Cref{fig:app:peristalsis}). Each node is assigned to the subnetwork
associated with the nearest supply node, and edges between nodes are
preserved (\Cref{fig:app:peristalsis}B). Next, we compute a minimal,
bifurcating and directed tree representation of each subnetwork:
$\mathcal{T}^1, \mathcal{T}^2, \mathcal{T}^3$
(\Cref{fig:app:peristalsis}C). Each tree $\mathcal{T}^j = \cup E_n^j$
consists of the subset of the nodes from $\Lambda^j$ that have degree
1 (are leaf or root nodes) or degree $3$ (are true bifurcation
points), and each path between nodes with degree $\not = 2$ in
$\Lambda^j$ is represented in $\mathcal{T}^j$ by an edge $E^j$ with
edge length $L$ corresponding to the total length of the original path
and edge radius $R^j$ as the average of the path radii.

For each subtree $\mathcal{T}^j = \cup_n E_n^j$, we compute the
time-averaged downstream flow rate $\langle Q'_n \rangle$ induced by
the vascular wall motion for each edge $n$ via~\cite[eq.~(5),
  (34)]{gjerde2023directional} (\Cref{fig:app:peristalsis}D). We next
assign this flow rate $\langle Q'_n \rangle$ to each of the branches
$\Lambda_i$ that form the path $E_n^j$, thus yielding $\langle Q'_i
\rangle$ for each perivascular segment $\Lambda_i$ while ensuring that
mass is conserved (\Cref{fig:app:peristalsis}E). For the segment(s)
ignored in the separation step (\Cref{fig:app:peristalsis}B), we set a
flow rate of zero. Finally, we define the mean longitudinal
perivascular velocity induced by the peristaltic wave $\langle
\hat{u}^x_i \rangle$ by dividing the flow rate $\langle Q'_i \rangle$
by the cross-section area $A_i = \pi ({R_2^i}^2 - {R_1^i}^2) = \pi
(\beta_i^2 - 1) {R_1^i}^2$.

@@

We assess the accuracy and numerical convergence of our simulation results by performing a series of experiments with different spatial and temporal resolutions. Specifically, we generate a sequence of three meshes, see \Cref{fig:mesh_refinement}, with an increasing number of points and cells, solve all relevant simulation steps (CSF flow and intracranial transport computations) for standard model A, and compare the results with respect to a set of key quantities of interest. The mesh refinement employed here is adaptive in the sense that the coarser mesh is refined near expected sharp concentration gradients. This allows us to better capture dynamics and reduce undershoots, see \Cref{fig:mesh_convergence_concentrations}.  
Similarly, we investigate the effect of the time step size by solving the intracranial transport model with varying time steps $dt \in \{60, 120, 240 \}$\,s.

@@

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/csf_minmax.png}
        \caption{Minimum and maximum tracer concentration in  the CSF under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1 \linewidth]{figures/par_minmax.png}
        \caption{Minimum and maximum tracer concentration in the parenchyma under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width = 1 \linewidth]{figures/art_minmax.png}
        \caption{Minimum and maximum tracer concentration in PVSs around arteries under mesh refinement}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
     \includegraphics[width= 1 \linewidth]{figures/ven_minmax.png}
         \caption{Minimum and maximum tracer concentration in PVSs around veins under mesh refinement}
    \end{subfigure}
    \caption{Minimum and maximum tracer concentrations over the first 24\,h after injection on the CSF and parenchyma (a), and the arterial and venous PVS (b) computed on the LowRes, standard and HighRes meshes with a timestep of $dt=120$\,s.}
    \label{fig:mesh_convergence_concentrations}
\end{figure}
